---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "# ",
  eval = TRUE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
Sys.setenv(PATH = paste(Sys.getenv("PATH"), path.expand("~/go/bin"), sep = ":"))
```

# glockr

*A **g**o **l**anguage c**o**de **c**ounter pac**k**age in **r***

<!-- badges: start -->
<!-- badges: end -->

The goal of `glockr` is to wrap the [`scc` code counter](https://github.com/boyter/scc) in an R package: 

> "*Goal is to be the fastest code counter possible, but also perform COCOMO calculation like sloccount, LOCOMO estimation for LLM-based development costs, estimate code complexity similar to cyclomatic complexity calculators and produce unique lines of code or DRYness metrics. In short one tool to rule them all.*"

This project was hugely inspired by the [`cloc` R package](https://github.com/hrbrmstr/cloc) by [boB Rudis](https://github.com/hrbrmstr).

## Installation

You can install the development version of `glockr` from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("mjfrigaard/glockr")
```

## Examples

Below is an example using the popular [`dplyr` package](https://dplyr.tidyverse.org/). 

```{r example}
library(glockr)
scc_version()
```

```{r}
pkg_path <- system.file(package = "dplyr")
```

```{r}
scc(pkg_path)
```

Filter to just the R source files and sort by `complexity` for a tighter view:

```{r}
r_files <- scc_by_file(pkg_path, include_ext = "r")
r_files_sub <- r_files[, c("filename", "lines", "code", "comments", "blanks", "complexity")]
r_files_sub[order(r_files_sub$complexity, decreasing = TRUE), ]
```

`scc` estimates cyclomatic complexity by counting control-flow keywords (`if`, `for`, `while`, etc.). The `complexity` column is the raw total; `weighted_complexity` divides by lines of code to give a per-line rate.

```{r}
r_files[order(-r_files$weighted_complexity),
        c("filename", "code", "complexity", "weighted_complexity")]
```

`scc_languages()` returns every language `scc` recognizes, along with the file extensions it maps to that language.

```{r}
langs <- scc_languages()
```

Number of languages:

```{r}
nrow(langs)
```

Also includes the extensions:

```{r}
head(langs, 10)
```


Last published: `r Sys.time()`