---
title: "Counting by File"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Counting by File}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 5
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path    <- glockr:::fetch_rlang_source()
if (is.na(rlang_path)) {
  message("rlang source unavailable; downstream chunks shown but not evaluated.")
  knitr::opts_chunk$set(eval = FALSE)
  rlang_path <- ""
}
```

```{r library}
library(glockr)
library(gt)
```

## Counting by file

`scc_by_file()` expands results to one row per source file, adding `filename`, `location`, and logical flags for `generated` and `minified` files.

```{r}
rlang_by_file <- scc_by_file(rlang_path) 
str(rlang_by_file)
```


Let's take a look at this by `complexity`, and drop `location`, `bytes`, `generated`, and `minified`. 

```{r rlang_by_file_cm}
rlang_by_file_cm <- rlang_by_file[ , c("language", "filename", "lines", 
                                        "blanks", "comments", "uloc", 
                                        "complexity", "weighted_complexity")]
rlang_by_file_cm[order(rlang_by_file_cm$complexity, decreasing = TRUE), ] |> 
  head(10) |> 
  gt::gt() |> 
  gt::tab_header(
    title = "Top 10 Complex Files", 
    subtitle = "by complexity", 
    preheader = "rlang package")
```

Filter to just the R source files and sort by complexity:

```{r scc-by-file-r}
r_files <- scc_by_file(rlang_path, include_ext = "r")
r_files_sub <- r_files[, c("filename", "lines", "code", "comments", "blanks", "complexity")]
r_files_sub[order(r_files_sub$complexity, decreasing = TRUE), ] |> 
  head(10) |> 
  gt::gt() |> 
  gt::tab_header(
    title = "Top 10 Complex R Files", 
    subtitle = "rlang package", 
    preheader = "rlang package")
```

## Visualizing file composition

With `ggplot2` and `tidyr` we can plot the code / comment / blank split for the most complex R source files.

```{r plot-pkgs, eval = scc_available}
library(ggplot2)
library(tidyr)
```

```{r tidytop15lines, eval = scc_available}
r_files <- scc_by_file(rlang_path, include_ext = "r")
# keep the 15 files with the highest lines of code
top15lines <- r_files[order(r_files$lines, decreasing = TRUE), ][1:15, ]
# tidy this
tidytop15lines <- tidyr::pivot_longer(
  data = top15lines,
  cols = c(code, comments, blanks),
  names_to = "type",
  values_to = "n_lines"
)
# change to factor
tidytop15lines$type <- factor(tidytop15lines$type, levels = c("blanks", "comments", "code"))
# keep only needed columns
tidytop15lines <- tidytop15lines[, c("n_lines", "type", "filename")]
str(tidytop15lines)
```

```{r}
ggplot2::ggplot(tidytop15lines, 
  ggplot2::aes(x = n_lines, y = reorder(filename, n_lines), fill = type)) +
  ggplot2::geom_col() +
  ggplot2::scale_fill_manual(
    values = c(code = "#2166ac", comments = "#92c5de", blanks = "#d1e5f0"),
    labels = c(code = "Code", comments = "Comments", blanks = "Blanks")
  ) +
  ggplot2::labs(
    title = "Stacked bar chart showing code, comment, and blank line counts for\nthe 15 largest R source files in rlang",
    x     = "Lines",
    y     = NULL,
    fill  = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")
```


```{r tidytop15cm, eval = scc_available}
# keep the 15 files with the highest lines of code
top15cm <- r_files[order(r_files$lines, decreasing = TRUE), ][1:15, ]
# tidy this
tidytop15cm <- tidyr::pivot_longer(
  data = top15cm,
  cols = c(complexity, weighted_complexity, uloc),
  names_to = "c_metric",
  values_to = "value"
)
# change to factor
tidytop15cm$c_metric <- factor(tidytop15cm$c_metric, levels = c("uloc", "complexity", "weighted_complexity"))
# keep only needed columns
tidytop15cm <- tidytop15cm[, c("lines", "filename", "c_metric", "value")]
str(tidytop15cm)
```

```{r plot-tidytop15cm, eval = scc_available}
ggplot2::ggplot(tidytop15cm, 
  ggplot2::aes(x = lines, y = reorder(filename, lines), fill = c_metric)) +
  ggplot2::geom_col() +
  ggplot2::scale_fill_manual(
    values = c(weighted_complexity = "#2166ac", complexity = "#92c5de", uloc = "#d1e5f0"),
    labels = c(weighted_complexity = "weighted complexity", complexity = "complexity", uloc = "uloc")
  ) + 
  ggplot2::labs(
    title = "Stacked bar chart showing ULOC, complexity, and weighted\ncomplexity for the 15 largest R source files in rlang",
    x     = "Lines",
    y     = NULL,
    fill  = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")
```