---
title: "COCOMO cost estimates"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{COCOMO cost estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "# ",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path    <- glockr:::fetch_rlang_source()
if (is.na(rlang_path)) {
  message("rlang source unavailable; downstream chunks shown but not evaluated.")
  knitr::opts_chunk$set(eval = FALSE)
  rlang_path <- ""
}
```

```{r library}
library(glockr)
library(gt)
```


## COCOMO cost estimates

`scc` includes a [COCOMO 81](https://en.wikipedia.org/wiki/COCOMO) model that estimates project effort and cost from `SLOC`. `scc` only emits COCOMO output via its tabular formatter — never in its JSON output — so `glockr` opts in by running `scc` a second time in tabular mode, scraping the COCOMO block, and parsing it into a tibble.

### Opting in with `cocomo = TRUE`

By default `scc()` returns the per-language tibble and ignores all COCOMO-related arguments. Pass `cocomo = TRUE` to also compute and return the COCOMO summary:

```{r cocomo}
rlang_cocomo <- scc(rlang_path, cocomo = TRUE)
```

The returned value is now a list (returned invisibly) with two tibbles:

```{r cocomo-shape}
names(rlang_cocomo)
```

```{r}
rlang_cocomo$cocomo |> 
  gt::gt() |> 
  gt::tab_header(
    title = "COCOMO Estimates", 
    subtitle = "rlang package", 
    preheader = "rlang package")
```

### Customize the COCOMO model

The customization arguments only take effect when `cocomo = TRUE`:

* `avg_wage`: annual salary in local currency   
* `cocomo_project_type`: "organic", "semi-detached", or "embedded"
* `eaf`: effort adjustment factor   
* `overhead`: corporate overhead multiplier   
* `currency_symbol`: symbol shown in the value column   
* `auto_print_scc`: skip auto-printing the language tibble      

```{r cocomo-embedded}
embedded <- scc(rlang_path,
    cocomo              = TRUE,          # required to compute COCOMO
    avg_wage            = 120000L,        
    cocomo_project_type = "embedded",     
    eaf                 = 1.1,            
    overhead            = 2.0,            
    currency_symbol     = "$",            
    auto_print_scc      = FALSE)         
```

```{r}
embedded$cocomo |> 
  gt::gt() |> 
  gt::tab_header(
    title = "COCOMO", 
    subtitle = "embedded project", 
    preheader = "rlang package")
```

```{r cocomo-organic}
organic <- scc(rlang_path,
    cocomo              = TRUE,
    avg_wage            = 140000L,
    cocomo_project_type = "organic",
    eaf                 = 1.1,
    overhead            = 2.0,
    currency_symbol     = "$",
    auto_print_scc      = FALSE)
```

```{r}
organic$cocomo |> 
  gt::gt() |> 
  gt::tab_header(
    title = "COCOMO", 
    subtitle = "organic project", 
    preheader = "rlang package")
```

### Quieting the auto-print

When `cocomo = TRUE`, `scc()` prints both tibbles to the console as a side effect. Set `auto_print_scc = FALSE` and/or `auto_print_cocomo = FALSE` to suppress either one and just keep the returned list:

```{r cocomo-silent}
silent <- scc(rlang_path,
              cocomo            = TRUE,
              auto_print_scc    = FALSE,
              auto_print_cocomo = FALSE)
```

### SLOCCount-compatible format

`sloccount_format = TRUE` reformats the upstream COCOMO block in SLOCCount-compatible form (useful for tooling that parses that output). 

```{r sloccount, comment="", collapse=FALSE}
sloccount <- scc(rlang_path,
                 cocomo           = TRUE,
                 sloccount_format = TRUE,
                 auto_print_scc   = FALSE)
```

The block doesn't fit the standard `metric / project_type / value` shape, so it's printed verbatim and `$cocomo` is `NULL`:

```{r}
is.null(sloccount$cocomo)
```

### Why not use COCOMO ???

`scc() `only gives us raw structural numbers (lines, code, complexity, bytes, etc.), but COCOMO II needs 22 judgement inputs we can't derive from source code:    

  * The current COCOMO project types (`"organic"` / `"semi-detached"` / `"embedded"`) don't translate to COCOMO  II    
  * COCOMO II uses 5 scale factors (PREC, FLEX, RESL, TEAM, PMAT) for organizational/process attributes and  17 effort multipliers (RELY, DATA, CPLX, RUSE, DOCU, TIME, STOR, PVOL, ACAP, PCAP, PCON, APEX, PLEX, LTEX, TOOL, SITE, SCED) on product, platform, personnel, and project context   
  * Every one of these is a human rating (VL /  L /  N / H / VH / XH) about the team, requirements stability, reuse strategy, tool maturity, schedule pressure, etc. 
  
The only one that *might*  map would be CPLX (product complexity), but COCOMO II's CPLX is architectural/algorithmic complexity, not  McCabe's cyclomatic.