---
title: "Getting Started with glockr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with glockr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
pkg_path <- system.file(package = "glockr")
```

`glockr` is an R wrapper around [scc](https://github.com/boyter/scc) (Succinct
Code Counter), a fast command-line tool that counts lines of code, comments, and
blank lines across hundreds of languages. `scc` also computes cyclomatic
complexity estimates, making it useful beyond simple line counting.

`scc` must be installed and on your `PATH`. See the
[scc releases page](https://github.com/boyter/scc/releases) for binaries.

## Functions

| Function | Description |
|---|---|
| `scc()` | Per-language summary for one or more paths |
| `scc_by_file()` | Per-file breakdown for one or more paths |
| `scc_languages()` | All languages recognized by `scc` |
| `scc_version()` | Installed `scc` version string |

## Setup

```{r library}
library(glockr)
```

```{r version, eval = scc_available}
scc_version()
```

## Counting by language

`scc()` returns one row per language. Here we run it against the `glockr`
package source itself.

```{r scc-pkg, eval = scc_available}
scc(pkg_path)
```

Results are sorted by `code` lines by default. Change the order with `sort`:

```{r scc-sort, eval = scc_available}
scc(pkg_path, sort = "files")
```

### Filtering by extension

Use `include_ext` to keep only certain file types. Extensions are
case-insensitive and match without the leading dot.

```{r scc-include, eval = scc_available}
scc(pkg_path, include_ext = "r")
```

`exclude_ext` works the opposite way — exclude one or more extensions:

```{r scc-exclude, eval = scc_available}
scc(pkg_path, exclude_ext = "r")
```

### Skipping complexity

Complexity calculation adds a small cost on large trees. Pass
`no_complexity = TRUE` to skip it:

```{r scc-no-complexity, eval = scc_available}
scc(pkg_path, no_complexity = TRUE)
```

## Counting by file

`scc_by_file()` expands results to one row per source file, adding `filename`,
`location`, and logical flags for `generated` and `minified` files.

```{r scc-by-file, eval = scc_available}
scc_by_file(pkg_path)
```

Filter to just the R source files for a tighter view:

```{r scc-by-file-r, eval = scc_available}
r_files <- scc_by_file(pkg_path, include_ext = "r")
r_files[, c("filename", "lines", "code", "comments", "blanks", "complexity")]
```

## Visualizing file composition

With `ggplot2` and `tidyr` we can plot the code / comment / blank split for
each source file.

```{r plot, eval = scc_available && requireNamespace("ggplot2", quietly = TRUE) && requireNamespace("tidyr", quietly = TRUE)}
library(ggplot2)
library(tidyr)

long <- pivot_longer(
  r_files,
  cols      = c(code, comments, blanks),
  names_to  = "type",
  values_to = "n_lines"
)
long$type <- factor(long$type, levels = c("blanks", "comments", "code"))

ggplot(long, aes(x = n_lines, y = reorder(filename, n_lines), fill = type)) +
  geom_col() +
  scale_fill_manual(
    values = c(code = "#2166ac", comments = "#92c5de", blanks = "#d1e5f0"),
    labels = c(code = "Code", comments = "Comments", blanks = "Blanks")
  ) +
  labs(
    title = "glockr R source — line composition",
    x     = "Lines",
    y     = NULL,
    fill  = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(legend.position = "bottom")
```

## Complexity metrics

`scc` estimates cyclomatic complexity by counting control-flow keywords
(`if`, `for`, `while`, etc.). The `complexity` column is the raw total;
`weighted_complexity` divides by lines of code to give a per-line rate.

```{r complexity, eval = scc_available}
r_files[order(-r_files$complexity),
        c("filename", "code", "complexity", "weighted_complexity")]
```

## Analysing multiple paths

Both `scc()` and `scc_by_file()` accept a character vector of paths, so you
can compare directories side by side:

```{r multi-path, eval = scc_available}
scc(c(
  file.path(pkg_path, "R"),
  file.path(pkg_path, "tests")
))
```

## Supported languages

`scc_languages()` returns every language `scc` recognizes, along with the file
extensions it maps to that language.

```{r languages, eval = scc_available}
langs <- scc_languages()
nrow(langs)

head(langs, 10)
```

Search for a specific language:

```{r lang-search, eval = scc_available}
langs[grepl("^R$", langs$language), ]
```
