---
title: "Getting Started with glockr"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Getting Started with glockr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path <- file.path("../inst/pkgs/rlang")
```

`glockr` is an R wrapper around [scc](https://github.com/boyter/scc) (Succinct
Code Counter), a fast command-line tool that counts lines of code, comments, and
blank lines across hundreds of languages. `scc` also computes cyclomatic
complexity estimates, making it useful beyond simple line counting.

`scc` must be installed and on your `PATH`. See the
[scc releases page](https://github.com/boyter/scc/releases) for binaries.

## Functions

| Function | Description |
|---|---|
| `scc()` | Per-language summary for one or more paths |
| `scc_by_file()` | Per-file breakdown for one or more paths |
| `scc_languages()` | All languages recognized by `scc` |
| `scc_version()` | Installed `scc` version string |

## Setup

```{r library}
library(glockr)
library(gt)
```

```{r version, eval = TRUE}
scc_version()
```

## Counting by language

`scc()` returns one row per language. The examples below use the `rlang` source
bundled at `inst/pkgs/rlang`.[^rlang-download]

[^rlang-download]: This is a downloaded a local version of the [source files](https://github.com/r-lib/rlang/) as of 2026-05-13.

```{r scc-pkg, eval = TRUE}
scc(rlang_path) |> 
  gt::gt(caption = "Succinct Code Counter (rlang R package)", 
    rowname_col = "language") 
```

Results are sorted by `code` lines by default. Change the order with `sort`:

```{r scc-sort, eval = TRUE}
scc(rlang_path, sort = "complexity") |> 
  gt::gt(caption = "Succinct Code Counter (sorted by complexity)", 
    rowname_col = "language") 
```

Valid `sort` values are: `"name"`, `"files"`, `"lines"`, `"code"`, `"comments"`, `"blanks"`, `"complexity"`.

### Filtering by extension

Use `include_ext` to keep only certain file types. Extensions are case-insensitive and match without the leading dot.

```{r scc-include, eval = TRUE}
scc(rlang_path, include_ext = "r") |> 
  gt::gt(caption = "Succinct Code Counter (by file extension)", 
  rowname_col = "language") 
```

`exclude_ext` works the opposite way — exclude one or more extensions:

```{r scc-exclude, eval = TRUE}
scc(rlang_path, exclude_ext = "r") |> 
  gt::gt(caption = "Succinct Code Counter (omit file extensions)", 
  rowname_col = "language") 
```

Pass a character vector to filter on multiple extensions at once:

```{r scc-multi-ext, eval = TRUE}
scc(rlang_path, include_ext = c("r", "c")) |> 
  gt::gt(caption = "Succinct Code Counter (.R and .c files)", 
         rowname_col = "language") 
```

### Skipping complexity

Complexity calculation adds a small cost on large trees. Pass
`no_complexity = TRUE` to skip it:

```{r scc-no-complexity, eval = TRUE}
scc(rlang_path, no_complexity = TRUE) |> 
  gt::gt(caption = "Succinct Code Counter (no complexity calculations)", 
         rowname_col = "language") 
```
