---
title: "Getting Started with glockr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with glockr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path <- system.file("pkgs/rlang", package = "glockr")
```

`glockr` is an R wrapper around [scc](https://github.com/boyter/scc) (Succinct
Code Counter), a fast command-line tool that counts lines of code, comments, and
blank lines across hundreds of languages. `scc` also computes cyclomatic
complexity estimates, making it useful beyond simple line counting.

`scc` must be installed and on your `PATH`. See the
[scc releases page](https://github.com/boyter/scc/releases) for binaries.

## Functions

| Function | Description |
|---|---|
| `scc()` | Per-language summary for one or more paths |
| `scc_by_file()` | Per-file breakdown for one or more paths |
| `scc_languages()` | All languages recognized by `scc` |
| `scc_version()` | Installed `scc` version string |

## Setup

```{r library}
library(glockr)
```

```{r version, eval = FALSE}
scc_version()
#> [1] "scc version 3.7.0"
```

## Counting by language

`scc()` returns one row per language. The examples below use the `rlang` source
bundled at `inst/pkgs/rlang`.

```{r scc-pkg, eval = FALSE}
scc(rlang_path)
#> # A tibble: 11 × 10
#>    language   files lines  code comments blanks
#>    <chr>      <int> <int> <int>    <int>  <int>
#>  1 R            163 44571 28038    11152   5381
#>  2 Markdown      53 16935 13866        0   3069
#>  3 C             70 14014 10960      872   2182
#>  4 C Header      72  8210  5315     1777   1118
#>  5 YAML           9   735   587       35    113
#>  6 C++            2    25    21        0      4
#>  7 C++ Header     1    26    21        0      5
#>  8 SVG           10    10    10        0      0
#>  9 Makefile       1    11     7        0      4
#> 10 License        1     2     2        0      0
#> 11 TOML           1     2     2        0      0
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

Results are sorted by `code` lines by default. Change the order with `sort`:

```{r scc-sort, eval = FALSE}
scc(rlang_path, sort = "files")
#> # A tibble: 11 × 10
#>    language   files lines  code comments blanks
#>    <chr>      <int> <int> <int>    <int>  <int>
#>  1 R            163 44571 28038    11152   5381
#>  2 C Header      72  8210  5315     1777   1118
#>  3 C             70 14014 10960      872   2182
#>  4 Markdown      53 16935 13866        0   3069
#>  5 SVG           10    10    10        0      0
#>  6 YAML           9   735   587       35    113
#>  7 C++            2    25    21        0      4
#>  8 C++ Header     1    26    21        0      5
#>  9 License        1     2     2        0      0
#> 10 Makefile       1    11     7        0      4
#> 11 TOML           1     2     2        0      0
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

Valid `sort` values are: `"name"`, `"files"`, `"lines"`, `"code"`,
`"comments"`, `"blanks"`, `"complexity"`.

### Filtering by extension

Use `include_ext` to keep only certain file types. Extensions are
case-insensitive and match without the leading dot.

```{r scc-include, eval = FALSE}
scc(rlang_path, include_ext = "r")
#> # A tibble: 1 × 10
#>   language files lines  code comments blanks
#>   <chr>    <int> <int> <int>    <int>  <int>
#> 1 R          163 44571 28038    11152   5381
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

`exclude_ext` works the opposite way — exclude one or more extensions:

```{r scc-exclude, eval = FALSE}
scc(rlang_path, exclude_ext = "r")
#> # A tibble: 10 × 10
#>    language   files lines  code comments blanks
#>    <chr>      <int> <int> <int>    <int>  <int>
#>  1 Markdown      53 16935 13866        0   3069
#>  2 C             70 14014 10960      872   2182
#>  3 C Header      72  8210  5315     1777   1118
#>  4 YAML           9   735   587       35    113
#>  5 C++            2    25    21        0      4
#>  6 C++ Header     1    26    21        0      5
#>  7 SVG           10    10    10        0      0
#>  8 Makefile       1    11     7        0      4
#>  9 License        1     2     2        0      0
#> 10 TOML           1     2     2        0      0
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

Pass a character vector to filter on multiple extensions at once:

```{r scc-multi-ext, eval = FALSE}
scc(rlang_path, include_ext = c("r", "c"))
#> # A tibble: 2 × 10
#>   language files lines  code comments blanks
#>   <chr>    <int> <int> <int>    <int>  <int>
#> 1 R          163 44571 28038    11152   5381
#> 2 C           70 14014 10960      872   2182
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

### Skipping complexity

Complexity calculation adds a small cost on large trees. Pass
`no_complexity = TRUE` to skip it:

```{r scc-no-complexity, eval = FALSE}
scc(rlang_path, no_complexity = TRUE)
#> # A tibble: 11 × 10
#>    language   files lines  code comments blanks
#>    <chr>      <int> <int> <int>    <int>  <int>
#>  1 R            163 44571 28038    11152   5381
#>  2 Markdown      53 16935 13866        0   3069
#>  3 C             70 14014 10960      872   2182
#>  4 C Header      72  8210  5315     1777   1118
#>  5 YAML           9   735   587       35    113
#>  6 C++            2    25    21        0      4
#>  7 C++ Header     1    26    21        0      5
#>  8 SVG           10    10    10        0      0
#>  9 Makefile       1    11     7        0      4
#> 10 License        1     2     2        0      0
#> 11 TOML           1     2     2        0      0
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```
