---
title: "Code Metrics"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Code Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path <- file.path("../inst/pkgs/rlang")
```

```{r library}
library(glockr)
library(gt)
```

## Complexity metrics

`scc` estimates cyclomatic complexity by counting control-flow keywords (`if`, `for`, `while`, etc.). The `complexity` column is the raw total.

### Weighted complexity 

`weighted_complexity` reports cyclomatic complexity *per 100 lines of code*, using the same formula `scc` applies in its wide tabular output:

```
weighted_complexity = (complexity / code) * 100      # 0 when code == 0
```

Note that `scc`'s JSON formatter does **not** populate `WeightedComplexity` (the field is emitted as `0` regardless of input), so `glockr` computes the value itself from `complexity` and `code`. For `scc_by_file()` this matches `scc`'s per-file value exactly; for `scc()` the same formula is applied to each language's aggregate totals.

```{r complexity, eval = TRUE}
r_files <- scc_by_file(rlang_path, include_ext = "r")
```

```{r}
r_files[order(-r_files$complexity),
        c("filename", "code", "complexity", "weighted_complexity")] |>
  head(10) |>
  gt::gt(caption = "Top 10 Complex files (.R files)",
    rowname_col = "filename")
```

## Unique lines of code

`uloc = TRUE` asks `scc` to count *unique* lines of code (ULOC) — the number of distinct non-blank, non-comment lines. 

```{r uloc, eval = TRUE}
scc(rlang_path, uloc = TRUE) |>
  gt::gt(caption = "Succinct Code Counter (with ULOC calculations)", 
         rowname_col = "language") 
```

`dryness = TRUE` also computes ULOC and additionally prints a DRYness score in the console (it has no extra effect on the returned tibble).

```{r dryness, eval = TRUE}
scc(rlang_path, dryness = TRUE) |>
  gt::gt(caption = "Succinct Code Counter (with dryness = TRUE)", 
         rowname_col = "language") 
```

**The `uloc` column is now populated — for example, R has 23,837 unique code lines out of 28,038 total, suggesting about 15 % duplication across the codebase.**

## Language remapping

Below are arguments for including unknown file types and languages. 

### Unknown-extension files

`count_as` maps file extensions to known `scc` languages using the format `"ext:language"` (or several mappings separated by commas). Below we include Quarto and RMarkdown as Markdown. 

```{r count-as, eval = TRUE}
scc(rlang_path, count_as = "qmd:Markdown,rmd:Markdown") |>
    gt::gt(caption = "Succinct Code Counter (including .Qmd/.Rmd)", 
           rowname_col = "language") 
```

### Remap by content

`remap_unknown` inspects only unrecognized files and assigns a language when a header string matches. 

```{r remap_unknown, eval = TRUE}
scc(rlang_path, remap_unknown = "# R script:R") |>
  gt::gt(caption = "Succinct Code Counter (remap unknown # R script:R)", 
       rowname_col = "language")
```

`remap_all` does the same for every file, overriding the extension-based detection.

```{r remap_all, eval = TRUE}
scc(rlang_path, remap_all = "# R script:R") |>
  gt::gt(caption = "Succinct Code Counter (remap all # R script:R)", 
       rowname_col = "language")
```
