---
title: "Code Metrics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Code Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path <- system.file("pkgs/rlang", package = "glockr")
```

```{r library}
library(glockr)
```

## Complexity metrics

`scc` estimates cyclomatic complexity by counting control-flow keywords
(`if`, `for`, `while`, etc.). The `complexity` column is the raw total.

`weighted_complexity` reports cyclomatic complexity *per 100 lines of code*,
using the same formula `scc` applies in its wide tabular output:

```
weighted_complexity = (complexity / code) * 100      # 0 when code == 0
```

Note that `scc`'s JSON formatter does **not** populate `WeightedComplexity`
(the field is emitted as `0` regardless of input), so `glockr` computes the
value itself from `complexity` and `code`. For `scc_by_file()` this matches
`scc`'s per-file value exactly; for `scc()` the same formula is applied to
each language's aggregate totals.

```{r complexity, eval = FALSE}
r_files <- scc_by_file(rlang_path, include_ext = "r")

r_files[order(-r_files$complexity),
        c("filename", "code", "complexity", "weighted_complexity")]
#> # A tibble: 163 × 4
#>    filename               code complexity weighted_complexity
#>    <chr>                 <int>      <int>               <dbl>
#>  1 cnd-abort.R             691        152                22.0
#>  2 standalone-cli.R        432        135                31.2
#>  3 trace.R                 807        134                16.6
#>  4 deparse.R               819        111                13.6
#>  5 standalone-vctrs.R      505        107                21.2
#>  6 call.R                  451         92                20.4
#>  7 standalone-obj-type.R   206         64                31.1
#>  8 cnd-entrace.R           213         61                28.6
#>  9 utils.R                 302         59                19.5
#> 10 expr.R                  157         52                33.1
#> # ℹ 153 more rows
```

## Unique lines of code

`uloc = TRUE` asks `scc` to count *unique* lines of code (ULOC) — the number
of distinct non-blank, non-comment lines. `dryness = TRUE` also computes ULOC
and additionally prints a DRYness score in the console (it has no extra effect
on the returned tibble).

```{r uloc, eval = FALSE}
scc(rlang_path, uloc = TRUE)
#> # A tibble: 11 × 10
#>    language   files lines  code comments blanks
#>    <chr>      <int> <int> <int>    <int>  <int>
#>  1 R            163 44571 28038    11152   5381
#>  2 Markdown      53 16935 13866        0   3069
#>  3 C             70 14014 10960      872   2182
#>  4 C Header      72  8210  5315     1777   1118
#>  5 YAML           9   735   587       35    113
#>  6 C++            2    25    21        0      4
#>  7 C++ Header     1    26    21        0      5
#>  8 SVG           10    10    10        0      0
#>  9 Makefile       1    11     7        0      4
#> 10 License        1     2     2        0      0
#> 11 TOML           1     2     2        0      0
#> # ℹ 4 more variables: complexity <int>,
#> #   weighted_complexity <dbl>, bytes <int>,
#> #   uloc <int>
```

The `uloc` column is now populated — for example, R has 23 837 unique code
lines out of 28 038 total, suggesting about 15 % duplication across the
codebase.

## Language remapping

### Unknown-extension files

`count_as` maps file extensions to known `scc` languages using the format
`"ext:language"` (or several mappings separated by commas):

```{r count-as, eval = FALSE}
scc(rlang_path, count_as = "qmd:Markdown,rmd:Markdown")
```

### Remap by content

`remap_unknown` inspects only unrecognised files and assigns a language when a
header string matches. `remap_all` does the same for every file, overriding
the extension-based detection.

```{r remap, eval = FALSE}
scc(rlang_path, remap_unknown = "# R script:R")
scc(rlang_path, remap_all     = "# R script:R")
```

## COCOMO cost estimates

`scc` includes a
[COCOMO 81](https://en.wikipedia.org/wiki/COCOMO) model that estimates project
effort and cost from SLOC. The COCOMO output is printed to the console (tabular
mode) but does **not** add columns to the returned tibble.

```{r cocomo, eval = FALSE}
# suppress COCOMO lines from console output
scc(rlang_path, no_cocomo = TRUE)

# customise the model
scc(rlang_path,
    avg_wage            = 120000L,       # annual salary in local currency
    cocomo_project_type = "embedded",    # "organic", "semi-detached", or "embedded"
    eaf                 = 1.1,           # effort adjustment factor
    overhead            = 2.0,           # corporate overhead multiplier
    currency_symbol     = "€")           # symbol shown in console output
```

`sloccount_format = TRUE` prints the COCOMO block in SLOCCount-compatible
format (useful for tooling that parses that output):

```{r sloccount, eval = FALSE}
scc(rlang_path, sloccount_format = TRUE)
```
