---
title: "Code Metrics"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Code Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "# ",
  fig.path = "../man/figures/",
  fig.width = 7,
  fig.height = 4
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path    <- glockr:::fetch_rlang_source()
if (is.na(rlang_path)) {
  message("rlang source unavailable; downstream chunks shown but not evaluated.")
  knitr::opts_chunk$set(eval = FALSE)
  rlang_path <- ""
}
```

```{r library}
library(glockr)
library(gt)
```

## Complexity metrics

`scc`'s `complexity` column is an *approximation* of [cyclomatic complexity](https://en.wikipedia.org/wiki/Cyclomatic_complexity) — not a true measurement. It's computed without parsing the code into an AST: while `scc` is already scanning each file for line classification, it also increments a per-file counter every time it sees a token that typically introduces a branch. The exact token set depends on the language; in Java, for example, each of `for`, `if`, `switch`, `while`, `else`, `||`, `&&`, `!=`, and `==` adds 1.

A few caveats from scc's own [README](https://github.com/boyter/scc/blob/master/README.md#complexity-estimates) worth carrying into any analysis:

* **Same-language comparisons only.** The counter is calculated by looking for the branch tokens of a particular language; languages with fewer such tokens (or that rely on recursion instead of loops) score lower without being intrinsically simpler. Use `complexity` to rank files / projects *within* the same language, not to compare R against C.
* **No AST, by design.** The trade-off is speed: the count is essentially free during the regular pass, but recursive methods or other non-branching constructs are not picked up.
* **Practical use.** The most common pattern is finding the most complex files in a project. With `glockr` that's:

```{r top-complex}
r_files <- scc_by_file(rlang_path, include_ext = "r")
r_files[order(-r_files$complexity),
        c("filename", "code", "complexity")] |>
  head(10) |>
  gt::gt() |>
  gt::tab_header(
    title    = "Top 10 most complex .R files",
    subtitle = "raw branch-token count",
    preheader = "rlang package")
```

The `complexity` column is the raw branch-token count per record.

### Weighted complexity 

`weighted_complexity` reports cyclomatic complexity *per 100 lines of code*, using the same formula `scc` applies in its wide tabular output:

```
weighted_complexity = (complexity / code) * 100      # 0 when code == 0
```

Note that `scc`'s JSON formatter does **not** populate `WeightedComplexity` (the field is emitted as `0` regardless of input), so `glockr` computes the value itself from `complexity` and `code`. For `scc_by_file()` this matches `scc`'s per-file value exactly; for `scc()` the same formula is applied to each language's aggregate totals.

Sorting the same files by `weighted_complexity` instead surfaces small-but-dense files that the raw count under-emphasises (a 30-line file with complexity = 11 will rank higher than a 300-line file with complexity = 15):

```{r weighted-complexity-top}
r_files[order(-r_files$weighted_complexity),
        c("filename", "code", "complexity", "weighted_complexity")] |>
  head(10) |>
  gt::gt() |>
  gt::tab_header(
    title    = "Top 10 by weighted complexity",
    subtitle = "complexity per 100 lines of code",
    preheader = "rlang package (.R files)")
```

## Unique lines of code

`uloc` (Unique Lines Of Code) is enabled by default — pass `uloc = FALSE` to skip the calculation. Despite the name, `scc`'s per-record ULOC is the count of unique **non-blank** lines, which means **comments count toward ULOC**. SLOC (the `code` column) is just the raw count of non-blank, non-comment lines and includes duplicates.

The clearest way to see the difference is two minimal fixtures bundled with the package:

```r
# inst/test/file/code.R           — pure code, 5 lines, two duplicate pairs
x <- 1
y <- 2
x <- 1
y <- 2
z <- 3
```

```r
# inst/test/file/code_w_comments.R — same code, plus three unique comments
# step 1
x <- 1
y <- 2
# step 2
x <- 1
y <- 2
# step 3
z <- 3
```

Running `scc_by_file()` on both:

```{r uloc-fixtures}
fixtures <- c(
  system.file("test/file/code.R",            package = "glockr"),
  system.file("test/file/code_w_comments.R", package = "glockr")
)
scc_by_file(fixtures, dryness = TRUE)[, c("filename", "lines", "code",
                                          "comments", "uloc", "dryness")] |>
  gt::gt() |>
  gt::tab_header(
    title    = "SLOC vs ULOC on minimal fixtures",
    subtitle = "both files contain the same 5 code lines",
    preheader = "inst/test/file/")
```

Two takeaways:

1. **`code.R`**: SLOC = 5 but ULOC = 3. The lines `x <- 1` and `y <- 2` each appear twice; the unique-line set is `{x <- 1, y <- 2, z <- 3}`.
2. **`code_w_comments.R`**: Same SLOC = 5, but ULOC = 6. The unique-non-blank set is `{# step 1, # step 2, # step 3, x <- 1, y <- 2, z <- 3}` — three distinct comments expand ULOC above SLOC.

So **ULOC can be either smaller or larger than SLOC** depending on the mix of duplication and unique comments.

### ULOC at scale

The same calculation aggregates across an entire codebase:

```{r uloc}
scc(rlang_path) |>
  gt::gt() |>
  gt::tab_header(
    title = "Succinct Code Counter",
    subtitle = "per-language ULOC",
    preheader = "rlang package")
```

### DRYness

The `dryness` column is **opt-in**: pass `dryness = TRUE` to either `scc()` or `scc_by_file()` and a `dryness` column appears in the returned tibble, computed locally as `uloc / lines` (matching `scc`'s `DRYness %` formula in its tabular output, applied per record instead of project-wide). Values close to 1 mean a file or language is mostly unique non-blank content; values closer to 0 mean heavy duplication, comments, or blanks. Since the formula needs `uloc`, passing `dryness = TRUE` also auto-promotes `uloc` to `TRUE` even if you explicitly set it to `FALSE`.

```{r dryness}
scc(rlang_path, dryness = TRUE)[, c("language", "lines", "code",
                                    "uloc", "dryness")] |>
  gt::gt() |>
  gt::tab_header(
    title = "Per-language DRYness",
    subtitle = "dryness = uloc / lines",
    preheader = "rlang package")
```

For rlang, the `R` language scores about 0.54 — roughly half of all physical R lines are unique non-blank content, with the remainder being either duplicates, blank lines, or repeated comment lines.

## Language remapping

Below are arguments for including unknown file types and languages. 

### Unknown-extension files

`count_as` maps file extensions to known `scc` languages using the format `"ext:language"` (or several mappings separated by commas). Below we include Quarto and RMarkdown as Markdown. 

```{r count-as}
scc(rlang_path, count_as = "qmd:Markdown,rmd:Markdown") |>
  gt::gt() |> 
  gt::tab_header(
    title = "Succinct Code Counter", 
    subtitle = "including .Qmd/.Rmd", 
    preheader = "rlang package")
```

### Remap by content

`remap_unknown` inspects only unrecognized files and assigns a language when a header string matches. 

```{r remap_unknown}
scc(rlang_path, remap_unknown = "# R script:R") |>
  gt::gt() |> 
  gt::tab_header(
    title = "Succinct Code Counter", 
    subtitle = "remap unknown # R script:R", 
    preheader = "rlang package")
```

`remap_all` does the same for every file, overriding the extension-based detection.

```{r remap_all}
scc(rlang_path, remap_all = "# R script:R") |>
  gt::gt() |> 
  gt::tab_header(
    title = "Succinct Code Counter", 
    subtitle = "remap all # R script:R", 
    preheader = "rlang package")
```
