---
title: "Visualization"
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
vignette: >
  %\VignetteIndexEntry{Visualization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "# ",
  fig.path  = "../man/figures/",
  fig.width = 7,
  fig.height = 5
)

scc_available <- nchar(Sys.which("scc")) > 0L
rlang_path    <- glockr:::fetch_rlang_source()
if (is.na(rlang_path)) {
  message("rlang source unavailable; downstream chunks shown but not evaluated.")
  knitr::opts_chunk$set(eval = FALSE)
  rlang_path <- ""
}
```

```{r library}
library(glockr)
library(ggplot2)
library(tidyr)
```

This vignette collects ways of plotting `scc_by_file()` output. Each section pulls the rlang source via `glockr:::fetch_rlang_source()` and filters to `.R` files.

## Code / comment / blank composition

Stacked bar of the line composition for the 15 largest R source files:

```{r tidytop15lines}
r_files    <- scc_by_file(rlang_path, include_ext = "r")
top15lines <- r_files[order(r_files$lines, decreasing = TRUE), ][1:15, ]

tidytop15lines <- tidyr::pivot_longer(
  data      = top15lines,
  cols      = c(code, comments, blanks),
  names_to  = "type",
  values_to = "n_lines"
)
tidytop15lines$type <- factor(
  tidytop15lines$type,
  levels = c("blanks", "comments", "code")
)
```

```{r plot-composition}
ggplot2::ggplot(tidytop15lines,
  ggplot2::aes(x = n_lines, y = reorder(filename, n_lines), fill = type)) +
  ggplot2::geom_col() +
  ggplot2::scale_fill_manual(
    values = c(code = "#2166ac", comments = "#92c5de", blanks = "#d1e5f0"),
    labels = c(code = "Code", comments = "Comments", blanks = "Blanks")
  ) +
  ggplot2::labs(
    title = "Line composition for the 15 largest R source files in rlang",
    x     = "Lines",
    y     = NULL,
    fill  = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")
```

## Complexity and ULOC by file

Same files, but now stacking the three per-file numeric metrics that scale with size:

```{r tidytop15cm}
top15cm <- r_files[order(r_files$lines, decreasing = TRUE), ][1:15, ]

tidytop15cm <- tidyr::pivot_longer(
  data      = top15cm,
  cols      = c(complexity, weighted_complexity, uloc),
  names_to  = "c_metric",
  values_to = "value"
)
tidytop15cm$c_metric <- factor(
  tidytop15cm$c_metric,
  levels = c("uloc", "complexity", "weighted_complexity")
)
```

```{r plot-tidytop15cm}
ggplot2::ggplot(tidytop15cm,
  ggplot2::aes(x = lines, y = reorder(filename, lines), fill = c_metric)) +
  ggplot2::geom_col() +
  ggplot2::scale_fill_manual(
    values = c(weighted_complexity = "#2166ac",
               complexity          = "#92c5de",
               uloc                = "#d1e5f0"),
    labels = c(weighted_complexity = "weighted complexity",
               complexity          = "complexity",
               uloc                = "uloc")
  ) +
  ggplot2::labs(
    title = "ULOC, complexity, and weighted complexity for the 15 largest R files",
    x     = "Lines",
    y     = NULL,
    fill  = NULL
  ) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(legend.position = "bottom")
```

## Complexity vs. DRYness as a circle pack

Each circle is one `.R` source file. The circle area is the file's `weighted_complexity` (complexity per 100 LOC), so files that pack more branching per line are larger. 

The fill color is the file's `dryness` (`uloc / lines`); darker hues mean more duplication relative to physical line count, lighter hues mean more unique non-blank content.

```{r r_files_dryness}
r_files <- scc_by_file(rlang_path, include_ext = "r", dryness = TRUE)
```

Drop files with zero weighted_complexity (nothing to size the circle by).

```{r drop-r_files_dryness}
r_files <- r_files[r_files$weighted_complexity > 0, ]
str(r_files[, c("filename", "weighted_complexity", "dryness")])
```

Compute the circle layout with `packcircles::circleProgressiveLayout()` (area-proportional sizing) and turn it into polygon vertices for `ggplot2`. 

Tag every polygon vertex with the filename, weighted complexity, and dryness it belongs to so the static and interactive plots can both render labels:

```{r circle-pack-data-ProgressiveLayout}
packing <- packcircles::circleProgressiveLayout(
  r_files$weighted_complexity, sizetype = "area"
)
packing$filename            <- r_files$filename
packing$weighted_complexity <- r_files$weighted_complexity
packing$dryness             <- r_files$dryness
```

```{r circle-pack-data-Vertices}
polygons <- packcircles::circleLayoutVertices(packing, npoints = 60)
# circleLayoutVertices() numbered the vertices by the row position of r_files 
# polygons$id is being used as a row index into r_files (like a positional join)
polygons$filename            <- r_files$filename[polygons$id]
polygons$weighted_complexity <- r_files$weighted_complexity[polygons$id]
polygons$dryness             <- r_files$dryness[polygons$id]
```

### Static — top-N labelled

To keep the figure readable, only the top 10 circles by weighted complexity get a text label. Smaller circles still render but stay anonymous; their info is recoverable from the interactive version below.

```{r circle-pack-static}
top_labels <- packing[order(-packing$weighted_complexity), ][seq_len(10), ]

ggplot2::ggplot(polygons,
                ggplot2::aes(x, y, group = id, fill = dryness)) +
  ggplot2::geom_polygon(colour = "white", linewidth = 0.2) +
  shadowtext::geom_shadowtext(
    data    = top_labels,
    mapping = ggplot2::aes(x = x, y = y, label = filename),
    inherit.aes = FALSE,
    size      = 3,
    colour    = "#2B2B2B",     # Eames Charcoal (theme `secondary`)
    bg.colour = "white",       # halo so dark text stays legible on dark fills
    bg.r      = 0.15,
    fontface  = "bold"
  ) +
  ggplot2::scale_fill_viridis_c(
    name    = "DRYness\n(uloc / lines)",
    option  = "viridis",
    limits  = c(0, 1)
  ) +
  ggplot2::coord_equal() +
  ggplot2::labs(
    title    = "rlang R files: complexity vs. DRYness",
    subtitle = "circle area = weighted complexity; fill = uloc / lines; top 10 labelled",
    caption  = "scc_by_file(..., dryness = TRUE)"
  ) +
  ggplot2::theme_void(base_size = 11) +
  ggplot2::theme(
    plot.title      = ggplot2::element_text(face = "bold"),
    legend.position = "right"
  )
```

### Interactive — hover every circle

Same layout, but rendered via [`ggiraph`](https://davidgohel.github.io/ggiraph/) so each circle is its own SVG element with a tooltip. Hover any circle to read the file's name, weighted complexity, and DRYness — including the small ones that the static plot doesn't label.

```{r circle-pack-interactive}
gg_interactive <- ggplot2::ggplot(polygons,
                                  ggplot2::aes(x, y, group = id, fill = dryness)) +
  ggiraph::geom_polygon_interactive(
    mapping = ggplot2::aes(
      tooltip = sprintf(
        "%s\nweighted_complexity: %.2f\ndryness: %.3f",
        filename, weighted_complexity, dryness
      ),
      data_id = id
    ),
    colour    = "white",
    linewidth = 0.2
  ) +
  ggplot2::scale_fill_viridis_c(
    name    = "DRYness\n(uloc / lines)",
    option  = "viridis",
    limits  = c(0, 1)
  ) +
  ggplot2::coord_equal() +
  ggplot2::labs(
    title    = "rlang R files (interactive)",
    subtitle = "hover any circle for filename, weighted complexity, and DRYness"
  ) +
  ggplot2::theme_void(base_size = 11) +
  ggplot2::theme(
    plot.title      = ggplot2::element_text(face = "bold"),
    legend.position = "right"
  )

ggiraph::girafe(
  ggobj         = gg_interactive,
  width_svg     = 7,
  height_svg    = 5,
  options       = list(
    ggiraph::opts_hover(css  = "stroke:#1B2A41; stroke-width:1.5px;"),
    ggiraph::opts_tooltip(opacity = 0.95)
  )
)
```

The largest circles are the highest-density branching files; comparing their hue tells you whether that complexity is concentrated in repetitive code (low DRYness, dark) or in mostly unique code (high DRYness, bright).
